---
title: "Integrated population model for Auke Creek coho salmon"
output:
  html_document:
    fig_caption: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r set_options, echo=FALSE, cache=TRUE}
options(width = 120)
```

***

[__Mark D. Scheuerell__](https://faculty.washington.edu/scheuerl/)  
_Fish Ecology Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, Seattle, WA, mark.scheuerell@noaa.gov_


__Eric R. Buhle__  
_Quantitative Science Inc., Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, Seattle, WA USA, eric.buhle@noaa.gov_

***

This is version `r paste0('0.',format(Sys.time(), '%y.%m.%d'))`.

Note that author list and order subject to change.

***

# Requirements
All analyses require the [R software](https://cran.r-project.org/) (v3.4+) for data retrieval, data processing, and summarizing model results, and the [Stan software](http://mc-stan.org/) (v2.17.0) for Hamiltonian Monte Carlo (HMC) sampling.

We also need a few packages that are not included with the base installation of R, so we begin by installing them (if necessary) and then loading them.

```{r load_pkgs, message=FALSE, warning=FALSE}
## for inference
if(!require("rstan")) {
  install.packages("rstan")
  library("rstan")
}
if(!require("loo")) {
  install.packages("loo")
  library("loo")
}
if(!require("salmonIPM")) {
  devtools::install_github("ebuhle/salmonIPM")
  library("salmonIPM")
}
## for data munging
if(!require("readr")) {
  install.packages("readr")
  library("readr")
}
if(!require("dplyr")) {
  install.packages("dplyr")
  library("dplyr")
}
## for output/plotting
if(!require("viridis")) {
  install.packages("viridis")
  library("viridis")
}
## for dir management
if(!require("here")) {
  install.packages("here")
  library("here")
}
```

# Load fish data

We begin by reading in the fish data file assembled in Phase I. Specifically, it is a __.csv__ file with the following columns:

 * `pop`: Numeric or character population ID.
 * `year`: Numeric variable giving the year the fish spawned (i.e., the brood year).
 * `A`: Spawning habitat size (either stream length or area).
 * `S_tot_obs`: Total number (not density) of wild and hatchery-origin spawners.
 * `n_age_minAge`...`n_age_maxAge`: Multiple columns of the observed age distribution, where minAge (maxAge) is the numeral total age in years of the youngest (oldest) spawners.
 * `n_W_obs`: Observed frequency of natural-origin spawners.
 * `n_H_obs`: Observed frequency of hatchery-origin spawners.
 * `fit_p_HOS`: Logical or 0/1 indicating if, for each row in the data, the model should estimate p_HOS > 0.
 * `F_rate`: Total harvest rate (proportion) of natural-origin fish.
 * `B_take_obs`: Number of adults taken for hatchery broodstock.
 
```{r get_data}
## directory
datadir <- here("data")
## filename
datafile <- dir(datadir)[grep("ipm", dir(datadir))]
## read data
fishdat <- read_csv(file.path(datadir, datafile))
```

These data contain some early years prior to the start of the escapement and age data, so we will trim to the data from 1994 onward.

```{r trim_years}
## trim early years
fishdat <- filter(fishdat, year >= 1994)
```


# Temporary fixes

At this point there are still a few missing pieces of necessary data, so here we make the following edits.

## Spawning area

We will set this to 1 so carrying capacity will be in units "fish" rather than "fish per length/area".

```{r set_area}
# set spawning area
fishdat$A <- rep(1, dim(fishdat)[1])
```

## Harvest rates

We are lacking harvest rates for 2014-2016, so we set them to 0.05 for all populations and years.

```{r set_harvest}
## remove years with no F estimates
fishdat <- filter(fishdat, !is.na(F_rate))
```


# Fit model

```{r fit_model, eval=TRUE}
fit_ipm <- salmonIPM(fishdat, model = "IPM",
                     chains = 3, iter = 1000, warmup = 500,
                     control = list(adapt_delta = 0.95, stepsize = 0.01, max_treedepth = 13))
```

