---
title: "Integrated population model for Auke Creek coho salmon"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    fig_caption: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    highlight: haddock
    toc: yes
    number_sections: true
    toc_depth: '3'
fontsize: 11pt
geometry: margin=1in
---

```{r options, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE, highlight = TRUE, comment = NA, 
                      dev = "png", dev.args = list(type = "cairo-png"), dpi = 200,
                      out.width = "70%", fig.align = "center")

library(here)
if(!require(captioner))
  devtools::install_github("adletaw/captioner")
library(captioner)
fig_nums <- captioner("Figure ", suffix = ": ", auto_space = FALSE, style = "b", style_prefix = TRUE)
library(kableExtra)
```
```{r width, include=FALSE}
options(width = 130)
```
```{r read_chunks, echo = FALSE}
knitr::read_chunk(here("analysis","auke_coho_ipm.R"))
```

***

[__Mark D. Scheuerell__](https://faculty.washington.edu/scheuerl/)  
_Fish Ecology Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, Seattle, WA, mark.scheuerell@noaa.gov_


__Eric R. Buhle__  
_Quantitative Science Inc., Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, Seattle, WA USA, eric.buhle@noaa.gov_

***
Note that author list and order subject to change.
***

# Requirements
All analyses require the [R software](https://cran.r-project.org/) (v3.4+) for data retrieval, data processing, and summarizing model results, and the [Stan software](http://mc-stan.org/) (v2.17.0) for Hamiltonian Monte Carlo (HMC) sampling.

We also need a few packages that are not included with the base installation of R, so we begin by installing them (if necessary) and then loading them.

```{r setup, message=FALSE, warning=FALSE}
```

# Load Data

We begin by reading in the fish data file assembled in Phase I. Specifically, it is a __.csv__ file with the following columns:

 * `pop`: Numeric or character population ID.
 * `year`: Numeric variable giving the year the fish spawned (i.e., the brood year).
 * `A`: Spawning habitat size (either stream length or area).
 * `S_tot_obs`: Total number (not density) of wild and hatchery-origin spawners.
 * `n_age_minAge`...`n_age_maxAge`: Multiple columns of the observed age distribution, where minAge (maxAge) is the numeral total age in years of the youngest (oldest) spawners.
 * `n_W_obs`: Observed frequency of natural-origin spawners.
 * `n_H_obs`: Observed frequency of hatchery-origin spawners.
 * `fit_p_HOS`: Logical or 0/1 indicating if, for each row in the data, the model should estimate p_HOS > 0.
 * `F_rate`: Total harvest rate (proportion) of natural-origin fish.
 * `B_take_obs`: Number of adults taken for hatchery broodstock.
 
```{r data}
```

# Fit Models

## Density-Independent

```{r fit_exp, eval = !exists("fit_exp")}
```
```{r print_exp}
```

## Beverton-Holt

```{r fit_BH, eval = !exists("fit_BH")}
```
```{r print_BH}
```

## Ricker

```{r fit_Ricker, eval = !exists("fit_Ricker")}
```
```{r print_Ricker}
```

## Model Selection

Model comparison based on LOO. Unhelpful because Pareto *k*s are too high. We will proceed with the Ricker model for now because it gives more biologically reasonable estimates, especially for intrinsic productivity.

```{r loo, warning = FALSE}
```

# Plots

Estimated spawner-smolt production curve, showing observations and states

```{r plot_SR_Ricker, echo=FALSE, fig.width=7, fig.height=7, out.width="60%"}
mod_name <- "fit_Ricker"
<<plot_SR>>
```

Posterior distributions of spawner-recruit parameters, with priors overlaid

```{r plot_SR_params_Ricker, echo=FALSE, fig.width=10, fig.height=5, out.width="80%"}
mod_name <- "fit_Ricker"
<<plot_SR_params>>
```

Time series of observed smolt and spawner abundance with estimated states

```{r plot_smolt_spawner_timeseries_Ricker, echo=FALSE, fig.width=7, fig.height=7}
mod_name <- "fit_Ricker"
<<plot_smolt_spawner_timeseries>>
```

Time series of smolt recruitment process error residuals and the proportion of smolts from each cohort that outmigrated at age 2.

```{r plot_smolt_recruitment_timeseries_Ricker, echo=FALSE, fig.width=7, fig.height=7}
mod_name <- "fit_Ricker"
<<plot_smolt_recruitment_timeseries>>
```

Time series of smolt-to-adult survival for 2- and 3-year-old smolts, and the proportion of each smolt age that return as jacks rather than overwintering at sea.

```{r plot_SAR_jack_timeseries_Ricker, echo=FALSE, fig.width=7, fig.height=7}
mod_name <- "fit_Ricker"
<<plot_SAR_jack_timeseries>>
```
